Draft abstract for Envision primary outcome

### 2024-04-05 Christine Chan/EnVision CF/katie_nacfc_2024.qmd
Update katie_nacfc_2024.qmd

```{r data cleaning}
# Import
df <- read.csv("./Christine Chan/EnVision CF/Data_Raw/EnvisionCF_DATA_2024-03-14_1325.csv",
  na.strings = ""
)
# Fix missing access group
df$redcap_data_access_group[is.na(df$redcap_data_access_group)] <- "iowa"
# Glucose columns
glucose <- paste0("timepoint_", c(0, 10, 30, 60, 90, 120, 150, 180), "_min")
# Get diagnosis
df$Diagnosis <- NA
df$Diagnosis[df[, glucose[1]] < 100 &
  (rowSums(df[, glucose[2:5]] < 200, na.rm = T) == rowSums(!is.na(df[, glucose[2:5]]))) &
  (rowSums(df[, glucose[6:8]] < 140, na.rm = T) == rowSums(!is.na(df[, glucose[6:8]])))] <- "NGT"
df$Diagnosis[df[, glucose[1]] >= 100] <- "IFG"
df$Diagnosis[rowSums(df[, glucose[2:5]] >= 200, na.rm = T) > 0] <- "INDET"
df$Diagnosis[rowSums(df[, glucose[6:8]] >= 140, na.rm = T) > 0] <- "IGF"
df$Diagnosis[df$timepoint_0_min >= 126 | df$timepoint_120_min >= 200] <- "CFRD"
df$Diagnosis <- factor(df$Diagnosis, levels = c("NGT", "IFG", "INDET", "IGF", "CFRD"))
# iAUCs
df$iAUC30 <- apply(df, 1, function(r) {
  y <- as.numeric(r[glucose[1:3]]) - as.numeric(r[glucose[1]])
  if (is.na(last(y))) {
    return(NA)
  } else {
    x <- c(0, 10, 30)[!is.na(y)]
    y <- y[!is.na(y)]
    auc <- trapz(x, y)
    return(auc)
  }
})
df$iAUC120 <- apply(df, 1, function(r) {
  y <- as.numeric(r[glucose[1:6]]) - as.numeric(r[glucose[1]])
  if (is.na(last(y))) {
    return(NA)
  } else {
    x <- c(0, 10, 30, 60, 90, 120)[!is.na(y)]
    y <- y[!is.na(y)]
    auc <- trapz(x, y)
    return(auc)
  }
})
df$iAUC180 <- apply(df, 1, function(r) {
  y <- as.numeric(r[glucose]) - as.numeric(r[glucose[1]])
  if (is.na(last(y))) {
    return(NA)
  } else {
    x <- c(0, 10, 30, 60, 90, 120, 150, 180)[!is.na(y)]
    y <- y[!is.na(y)]
    auc <- trapz(x, y)
    return(auc)
  }
})
# Other formatting
df$date_visit <- lubridate::ymd(df$date_visit)
# Write for checking
df$Keep <- NULL
write.csv(df,
  file = "./Christine Chan/EnVision CF/Data_Clean/nacfc_2024_analysis_data.csv",
  row.names = F, na = ""
)
# Exclude those missing <= 3 missing OGTT values
df$Keep <- rowSums(is.na(df[, glucose])) <= 3
df <- df %>% filter(Keep)
```

@@

```{r}
#| label: tbl-2
#| tbl-cap: Participant diagnoses based on OGTT
df %>%
  select(
    redcap_data_access_group, Diagnosis, timepoint_0_min, timepoint_30_min,
    timepoint_60_min, timepoint_120_min, timepoint_180_min, iAUC30, iAUC120,
    iAUC180
  ) %>%
  tbl_summary(by = redcap_data_access_group, missing_text = "Missing") %>%
  add_overall()
```

@@

```{r}
#| label: tbl-3
#| tbl-cap: Participant diagnoses based on OGTT (visit 1 only)
df %>%
  group_by(study_id) %>%
  filter(row_number() == 1) %>%
  ungroup() %>%
  select(
    redcap_data_access_group, Diagnosis, timepoint_0_min, timepoint_30_min,
    timepoint_60_min, timepoint_120_min, timepoint_180_min, iAUC30, iAUC120,
    iAUC180
  ) %>%
  tbl_summary(by = redcap_data_access_group, missing_text = "Missing") %>%
  add_overall()
```

>>>>

```{r data cleaning}
# Import
df <- read.csv("./Christine Chan/EnVision CF/Data_Raw/EnvisionCF_DATA_2024-03-14_1325.csv",
  na.strings = ""
)
# Fix missing access group
df$redcap_data_access_group[is.na(df$redcap_data_access_group)] <- "iowa"
# Glucose columns
glucose <- paste0("timepoint_", c(0, 10, 30, 60, 90, 120, 150, 180), "_min")
# Get diagnosis
df$Diagnosis <- NA
df$Diagnosis[df[, glucose[1]] < 100 &
  (rowSums(df[, glucose[2:5]] < 200, na.rm = T) == rowSums(!is.na(df[, glucose[2:5]]))) &
  (rowSums(df[, glucose[6:8]] < 140, na.rm = T) == rowSums(!is.na(df[, glucose[6:8]])))] <- "NGT"
df$Diagnosis[df[, glucose[1]] >= 100] <- "IFG"
df$Diagnosis[rowSums(df[, glucose[2:5]] >= 200, na.rm = T) > 0] <- "INDET"
df$Diagnosis[rowSums(df[, glucose[6:8]] >= 140, na.rm = T) > 0] <- "IGF"
df$Diagnosis[df$timepoint_0_min >= 126 | df$timepoint_120_min >= 200] <- "CFRD"
df$Diagnosis <- factor(df$Diagnosis, levels = c("NGT", "IFG", "INDET", "IGF", "CFRD"))
# Add insulin data
insulin <- read.csv("./Christine Chan/EnVision CF/Data_Clean/insulin_wide.csv")
df <- left_join(df, insulin, by = join_by(study_id, date_visit))
insulin <- gsub("timepoint_", "insulin_", glucose)
# iAUCs
df$iAUC30gluc <- apply(df, 1, function(r) {
  y <- as.numeric(r[glucose[1:3]]) - as.numeric(r[glucose[1]])
  if (is.na(last(y))) {
    return(NA)
  } else {
    x <- c(0, 10, 30)[!is.na(y)]
    y <- y[!is.na(y)]
    auc <- trapz(x, y)
    return(auc)
  }
})
df$iAUC120gluc <- apply(df, 1, function(r) {
  y <- as.numeric(r[glucose[1:6]]) - as.numeric(r[glucose[1]])
  if (is.na(last(y))) {
    return(NA)
  } else {
    x <- c(0, 10, 30, 60, 90, 120)[!is.na(y)]
    y <- y[!is.na(y)]
    auc <- trapz(x, y)
    return(auc)
  }
})
df$iAUC180gluc <- apply(df, 1, function(r) {
  y <- as.numeric(r[glucose]) - as.numeric(r[glucose[1]])
  if (is.na(last(y))) {
    return(NA)
  } else {
    x <- c(0, 10, 30, 60, 90, 120, 150, 180)[!is.na(y)]
    y <- y[!is.na(y)]
    auc <- trapz(x, y)
    return(auc)
  }
})
df$iAUC30ins <- apply(df, 1, function(r) {
  y <- as.numeric(r[insulin[1:3]]) - as.numeric(r[insulin[1]])
  if (is.na(last(y))) {
    return(NA)
  } else {
    x <- c(0, 10, 30)[!is.na(y)]
    y <- y[!is.na(y)]
    auc <- trapz(x, y)
    return(auc)
  }
})
df$iAUC120ins <- apply(df, 1, function(r) {
  y <- as.numeric(r[insulin[1:6]]) - as.numeric(r[insulin[1]])
  if (is.na(last(y))) {
    return(NA)
  } else {
    x <- c(0, 10, 30, 60, 90, 120)[!is.na(y)]
    y <- y[!is.na(y)]
    auc <- trapz(x, y)
    return(auc)
  }
})
df$iAUC180ins <- apply(df, 1, function(r) {
  y <- as.numeric(r[insulin]) - as.numeric(r[insulin[1]])
  if (is.na(last(y))) {
    return(NA)
  } else {
    x <- c(0, 10, 30, 60, 90, 120, 150, 180)[!is.na(y)]
    y <- y[!is.na(y)]
    auc <- trapz(x, y)
    return(auc)
  }
})
# Hypoglycemia
df$Hypo70 <- factor(rowSums(df[, glucose] < 70, na.rm = T) > 0,
  levels = c(F, T), labels = c("No", "Yes")
)
df$Hypo60 <- factor(rowSums(df[, glucose] < 60, na.rm = T) > 0,
  levels = c(F, T), labels = c("No", "Yes")
)
df$Hypo50 <- factor(rowSums(df[, glucose] <= 50, na.rm = T) > 0,
  levels = c(F, T), labels = c("No", "Yes")
)
# HOMA IR
df$homa_ir = (df$timepoint_0_min * df$insulin_0_min)/405
# Write for Amir's abstract
write.csv(df,
  file = "./Christine Chan/EnVision CF/Data_Clean/nacfc_2024_analysis_data.csv",
  row.names = F, na = ""
)
# Exclude those missing <= 3 missing OGTT values
df$Keep <- NULL
df$Keep <- rowSums(is.na(df[, glucose])) <= 3
df <- df %>% filter(Keep)
```

@@

```{r}
#| label: tbl-2
#| tbl-cap: Participant diagnoses based on OGTT
df %>%
  select(
    redcap_data_access_group, Diagnosis, timepoint_0_min, timepoint_30_min,
    timepoint_60_min, timepoint_120_min, timepoint_180_min, 
    iAUC30gluc:iAUC180ins,homa_ir
  ) %>%
  tbl_summary(by = redcap_data_access_group, missing_text = "Missing") %>%
  add_overall()
```

@@

```{r}
#| label: tbl-3
#| tbl-cap: Participant diagnoses based on OGTT (visit 1 only)
df %>%
  group_by(study_id) %>%
  filter(row_number() == 1) %>%
  ungroup() %>%
  select(
    redcap_data_access_group, Diagnosis, timepoint_0_min, timepoint_30_min,
    timepoint_60_min, timepoint_120_min, timepoint_180_min, 
    iAUC30gluc:iAUC180ins,homa_ir
  ) %>%
  tbl_summary(by = redcap_data_access_group, missing_text = "Missing") %>%
  add_overall()
```
